\chapter{Acknowledgments}

First, I would like to express my sincere appreciation to my advisors: Esther Pacitti, Patrick Valduriez and Marta Mattoso for their patient, helpful, effective guidance. They gave me excellent advice to improve my writing skills, helped me to develop my own ideas and corrected my papers, posters, articles. In addition, they offered me many opportunities to connect with reputable experts in different disciplines. 

I would like to also thank my family: Jirong Liu and Xiuqin Hao, for giving birth to me and providing me with continuous encouragement and support. You make me strong in heart so that I could pursue my dreams without hesitation.

I am also deeply grateful to my committee members, Pascal Molli and Sébastien Monnet and my evaluator Christine Morin for spending their time to evaluate my thesis.

% and help me to improve it by valuable feedback.

I would like to express my gratitude towards my internship supervisors and collaborators at Microsoft, Seattle, USA: Vinod Nair, Markus Weimer, Sriram Rao and Raghu Ramakrishnan.
I am thankful to the collaborators and contributors during my thesis. To Gabriel Antoniu, Alexandru Costan, Luis Pineda-Morales, Radu Tudoran, Raghu Ramakrishnan, Olivier Nano, Laurent Massoulie and Pierre-Louis Xech for the great collaboration that we had in the Z-CloudFlow project. 
To Vítor Silva, Daniel de Oliveira and Kary Ocaña for our productive collaboration in the MUSIC project.
I would also like to thank Weiwei Chen, who helped me with the experiments and gave me valuable advice on my ideas. 
Special thanks go to Idy Camara, Leye Wang, Qiu Han, Yue Li, Haoyi Xiong and Xin Jin, who offered me their advice on my research and my career planning. 
Thanks also go to Amelie Chi Zhou, who shared her valuable research experience with me.

I would also like to record my sincere thanks to all the current and former members of the Zenith team. Special thanks to Reza Akbarinia, Maximilien Servajean, Miguel Liroz, Mohamed Reda Bouadjenek, Saber Salah, Oleksandra Levchenko and Rim Moussa, who gave me helpful advice on my research work. 
Thanks also go to Tristan Allard, Florent Masseglia, Daniel Gaspar, Djamel-Edine Yagoubi, Sen Wang and Carlyna Bondiombouy, with whom, I often recall the happy time we spent together.

I am grateful to my friends who are particular helpful along this path. Thanks go to Zhan Li, Wenlong Liu and Liuyang Zhang for all the fun we had and their help on my thesis.

Finally, I would like to thank everyone at UM, LIRMM, MSR-INRIA, INRIA and all the other people who had a direct or indirect contribution to this work and were not mentioned above. I appreciate their help and support.